---
title: "A03_G34801183"
author: "Bisman Sachar"
date: "February 2, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Downloading data directly from the AWS link to the R mardown file and saving it in fkicks dataframe object

```{r}
fkicks <- read.table("https://s3.amazonaws.com/programmingforanalytics/footballkicks.csv", header = TRUE, sep = ",")

```


# PART A - Predicting Melvin's attemps at field goals using simple probabilities

### Analyzing the data using str() and summary()

```{r}
str(fkicks)
summary(fkicks)
```

It tells us that overall, Melvin scored 79.72% (759 out of 952 kicks) of the times he attempted. He played 90 matches & 862 practice matches. He attempted from 18 to 62 yards. yards is a continuous variable, goal is a discrete and categorical variable, practiceormatch is discrete & categorical variable

### Checking confusion matrix at Melvin's attempts using prop.table(). 

```{r}
prop.table(table(fkicks$goal, fkicks$yards))
prop.table(table(fkicks$goal, fkicks$practiceormatch))
```


### To get probabilities, we need to convert goal non-numeric levels into numeric levels

```{r}
fkicks$goal <- ifelse(fkicks$goal == "Y", 1,0)
```

### Calculating goal success rate differentiated by practice or match

```{r}
library(dplyr)
fkicks %>% summarise(avggoal = mean(fkicks$goal))
```
#### Melvin's overall success rate is 79.72%

```{r}
fkicks %>% group_by(practiceormatch) %>% summarise(avggoal = mean(goal)) 
```
#### Melvin's goal success rate in matches is 82.2%. Melvin's goal success rate in practice is 79.5%. 



# PART B - Predicting probabilites using logistic function.

a)
### i)Writing logistic function for practice

### Creating a subset of data with only Practice (P)
```{r}
fkicks1 <- fkicks %>% filter(practiceormatch=="P")
```

### Running logistic function and finding intercept values

```{r}
goalp.log = glm(goal ~ yards,
                  data = fkicks1,
                  family=binomial)
summary(goalp.log)
```

#### This gives us the logit function - ln(p/1-p) = 5.802 - 0.113 * yards, where p is the probability of goal

### ii) Writing logistic function for matches

### Creating a subset of data with only Match (M)
```{r}
fkicks2 <- fkicks %>% filter(practiceormatch=="M")
```

### Running logistic function and finding intercept values

```{r}
goalm.log = glm(goal ~ yards,
                  data = fkicks2,
                  family=binomial)
summary(goalm.log)
```

#### This gives us the logit function - ln(p/1-p) = 4.837 - 0.0812 * yards, where p is the probability of goal

### b) Finding probability of scoring a goal when Melvin kicks from 20, 40 and 60 yards in practice

### Creating a new pdata variable for yards values - 20, 40 & 60
```{r}
pdata <- c(20, 40, 60)
```

### Predicting values using predict() by passing the above pdata variable
```{r}
predict(goalp.log, data.frame(yards=pdata), type="response")
```

#### Melvin's probability of goal success for 20, 40 and 60 yards is 0.972, 0.779 and 0.268 respectively.

### c) Finding probability scoring a goal when Melvin kicks from 20, 40 and 60 yards in match

### Creating a new mdata variable for yards values - 20, 40 & 60
```{r}
mdata <- c(20, 40, 60)
```

### Predicting values using predict() by passing the above pdata variable
```{r}
predict(goalm.log, data.frame(yards=mdata), type="response")
```
#### Melvin's probability of goal success for 20, 40 and 60 yards is 0.961, 0.830 and 0.490 respectively.


# PART C - Plotting the logistic model using ggplot

```{r}
# Calling the ggplot2 library
library(ggplot2)

#Using fkicks data set in ggplot function, gave x variable as yards and y variable as goal
ggplot(fkicks, aes(x=yards, y=goal, color=practiceormatch)) +
 
# Using stat_smooth function, used glm - binomial (logit) model  
stat_smooth(method="glm", se=FALSE, method.args = list(family=binomial), size = 2) + 

# Gave labels to x & y axis, also gave the title to the plot
ylab("Proabability of scoring a goal") + xlab("Distance to goal in yards") + ggtitle("Proabability of scoring a goal")


```

